---
title: "App to predict species"
author: "TM"
date: "November 19, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction {.smaller}

- This application utilizes machine learning algoithms to predict the species of a plant based on the legnth and width of the petal and Sepal
- Data is based on the standard R dataset, IRIS 

### User guide  

- User the sliders to enter length and width for Sepal and Petal
- Prediction appears on the right hand side
- Code can be found on git hub https://github.com/tristanmaw/Shiny-app-for-predicting-species

### libraries used
```{r, warning= FALSE, error = FALSE, message = FALSE, echo = TRUE} 
library(randomForest)
library(scales)
library(shiny)
```

## Model build using Random Forest {.smaller}
```{r, warning= FALSE, error = FALSE, message = FALSE, echo = TRUE} 
set.seed(100)
df <- iris
n <- nrow(df)
sample_size <- round(n * 0.5,0)
training <- df[sample(n),][1:sample_size,]
testing <- df[sample(n),][(sample_size + 1):n,]
modfit <- randomForest(Species ~., data = training)
# Validation
pred <- predict(modfit,testing)
testing$predRight <- pred == testing$Species    ##when prection is correct
table(pred,testing$Species)   ## confusion matrix
percent(sum(testing$predRight) / nrow(testing))   ## accuracy calc
```
Accuracy is `r percent(sum(testing$predRight) / nrow(testing))`


## Shiny development - UI code {.smaller}
- Code can also be found on github https://github.com/tristanmaw/Shiny-app-for-predicting-species
```{r, warning= FALSE, error = FALSE, message = FALSE, eval = FALSE, echo = TRUE}
shinyUI(fluidPage(
  titlePanel("What species is this plant?"),
  sidebarLayout(
    sidebarPanel(
      sliderInput('SepalLength', 'Sepal Length',value = 5, min = 4, max = 8, step = 0.1),
      sliderInput('SepalWidth',  'Sepal Wigth',value = 3, min = 1, max = 5, step = 0.1),
      sliderInput('PetalLength', 'Petal Length',value = 5, min = 1, max = 7, step = 0.1),
      sliderInput('PetalWidth', 'Petal Width',value = 1, min = 0.1, max = 3, step = 0.1)
      #submitButton("Submit")
    ),
    mainPanel(
      h3("The predicted species is"),
      verbatimTextOutput("prediction")
      
    )
  )
))
```

## Shiny development - Server code  {.smaller}
```{r, warning= FALSE, error = FALSE, message = FALSE, eval = FALSE, echo = TRUE}
shinyServer(function(input, output){
  set.seed(100)
  df <- iris
  n <- nrow(df)
  sample_size <- round(n * 0.75,0)
    training <- df[sample(n),][1:sample_size,]
    modfit <- randomForest(Species ~., data = training)
    model_pred <- reactive({
    predict(modfit, newdata = data.frame(Sepal.Length = input$SepalLength, 
                                         Sepal.Width = input$SepalWidth, 
                                         Petal.Length = input$PetalLength, 
                                         Petal.Width = input$PetalWidth))
  })
  model_pred2 <- reactive({
    x <- as.data.frame(model_pred())
    rownames(x) <- ""
    colnames(x) <- ""
    x
  })
      output$prediction <- renderPrint({
    model_pred2()
  })
}
)
```




